Safe Nuclear Power: Instrumentation, Human Oversight, and Infrastructure Transition

### 2025-06-07 Slides/PaperTwoPresentation/presentation.tex
update presentation

% ECE4900W Communicating Engineering Solutions
% Author: Arturo Salinas-Aguayo <artjsalina5@uconn.edu>
\documentclass{beamer}
\usetheme{Warsaw}
\usecolortheme{dolphin}
\usefonttheme{professionalfonts}
\setbeamertemplate{navigation symbols}{}
\usepackage{graphicx}
\usepackage{tavox}

@@

\graphicspath{{../../images/}}
% Bibliography
\addbibresource{references.bib}

@@

\begin{frame}
  \titlepage
  \speak{Welcome. This presentation explores the technological, ethical, and infrastructural dimensions of nuclear power in the modern world.}
\end{frame}

\begin{frame}{Outline}
  \tableofcontents[pausesections]
  \speak{Here is an overview of the major themes we will discuss today.}
\end{frame}

@@

\begin{frame}{Motivation}
  \speak{Nuclear energy offers unmatched energy density and reliability. Yet its history demands vigilance, ethics, and technical excellence.}
  \begin{itemize}
    \item Low emissions, high stability
    \item Risk perception shaped by major accidents
    \item Focus of this work: instrumentation, human oversight, infrastructure
  \end{itemize}
\end{frame}

@@

\begin{frame}{Types of Reactors}
  \speak{Different reactor types use different coolants and designs, influencing their control strategies.}
  \begin{itemize}
    \item Pressurized Water Reactor (PWR)
    \item Boiling Water Reactor (BWR)
    \item CANDU (Heavy Water Reactor)
    \item Gas-Cooled Reactor (AGR)
  \end{itemize}
\end{frame}

\begin{frame}{Instrumentation in a PWR}
  \speak{This diagram shows how sensors monitor power, pressure, and temperature throughout a PWR.}
  \includegraphics[width=0.85\textwidth]{images/instrumentation.png}
\end{frame}

@@

\begin{frame}{SL-1 and Three Mile Island}
  \speak{SL-1 and Three Mile Island showed the risks of inadequate interlocks and overwhelming operator interfaces.}
  \begin{itemize}
    \item SL-1: control rod ejection, fatal steam explosion
    \item TMI: misread signals led to core exposure
    \item Both: highlight flaws in human-system integration
  \end{itemize}
\end{frame}

\begin{frame}{Chernobyl and Fukushima}
  \speak{Chernobyl and Fukushima illustrate the danger of flawed assumptions and the limits of automation.}
  \includegraphics[width=0.75\textwidth]{images/chernobylafter.jpg}
  \includegraphics[width=0.75\textwidth]{images/fukushimaafter.jpg}
\end{frame}

@@

\begin{frame}{Human Factors Engineering}
  \speak{Human factors engineering is vital in designing systems that support, not hinder, decision making.}
  \begin{itemize}
    \item Ergonomic control rooms
    \item Full-scope simulation
    \item Situational awareness tools
  \end{itemize}
\end{frame}

\begin{frame}{Ethics of Automation}
  \speak{Automation must be supervised. Tunnel vision and outdated assumptions can cause failure without human oversight.}
  \begin{itemize}
    \item Humans mitigate uncertainty and change
    \item Ethics demand human-in-the-loop
  \end{itemize}
\end{frame}

@@

\begin{frame}{Waste and Infrastructure}
  \speak{The economic burden of waste management and infrastructure decline limits nuclear's future.}
  \begin{itemize}
    \item High refueling and decommissioning costs
    \item No permanent U.S. waste site
    \item Industry contraction post-Fukushima
  \end{itemize}
\end{frame}

\begin{frame}{Hopeful Developments}
  \speak{New efforts like Vogtle and SMRs show a path forward. Nuclear remains vital for base-load generation.}
  \begin{itemize}
    \item Vogtle Unit 4 completed
    \item Passive safety designs
    \item Small Modular Reactors
  \end{itemize}
\end{frame}

@@

\begin{frame}{Final Thoughts}
  \speak{With humility and innovation, nuclear energy can be safe, ethical, and essential in the fight against climate change.}
  \begin{itemize}
    \item Ethics + engineering = resilient systems
    \item Human vigilance must never be replaced
    \item The atom, split with care, still holds promise
  \end{itemize}
\end{frame}

>>>>

% ECE4900W Communicating Engineering Solutions
% Author: Arturo Salinas-Aguayo <artjsalina5@uconn.edu>
\documentclass{beamer}
\usetheme{Warsaw}
\usecolortheme{dolphin}
\usefonttheme{professionalfonts}
\setbeamertemplate{navigation symbols}{}
\usepackage{graphicx}
\usepackage{biblatex}
\addbibresource{references.bib}
\graphicspath{{../../images/}}

@@

@@

\begin{frame}
  \titlepage
  \speak{Welcome. This presentation explores the technological, ethical, and infrastructural dimensions of nuclear power in the modern world. We will cover lessons from historic accidents, instrumentation, human-machine interface design, and current innovations. Each story reveals both vulnerability and resilience.}
\end{frame}

\begin{frame}{Outline}
  \tableofcontents[pausesections]
  \speak{Here is the roadmap for this presentation. Each section builds on the last, culminating in a call for renewed responsibility and innovation in the nuclear field.}
\end{frame}

@@

\begin{frame}{Motivation}
  \subsubsection*{Why Nuclear Now?}
  \speak{Nuclear energy provides unmatched energy density and emissions-free base-load power. However, public trust and sustainable deployment depend on ethical design, robust control systems, and trained human oversight. While wind and solar depend on weather and batteries, nuclear runs steadily. But one accident can erode decades of progress.}
  \begin{itemize}
    \item Zero-emission base-load power
    \item High energy return on investment
    \item Historical trauma shapes perception \autocite{nrc_chernobyl, worldnuclear_fukushima_review}
    \item Paper focuses on safety through instrumentation and oversight
  \end{itemize}
\end{frame}

@@

\begin{frame}{Types of Reactors}
  \subsubsection*{Diverse Architectures}
  \speak{Let’s begin with an overview of reactor types, since their thermal-hydraulic behavior and control logic affect how instrumentation is designed. Each has different challenges.}
  \begin{itemize}
    \item Pressurized Water Reactor (PWR): most common, used in submarines and civilian fleets
    \item Boiling Water Reactor (BWR): direct steam path to turbine
    \item CANDU: heavy water as moderator and coolant
    \item Gas-cooled (AGR): high outlet temperature for efficiency
  \end{itemize}
\end{frame}

\begin{frame}{Instrumentation in a PWR}
  \subsubsection*{Sensor Coverage}
  \speak{This diagram shows typical sensor and actuator placements in a PWR. Note the neutron flux monitors, hot leg thermocouples, pressurizer level gauges, and scram breakers. Each of these is essential for safe operation.}
  \includegraphics[width=0.85\textwidth]{instrumentation.png}
\end{frame}

@@

\begin{frame}{SL-1: Prompt Critical from Manual Control}
  \subsubsection*{The First Fatal Reactor Accident}
  \includegraphics[width=0.4\textwidth]{sl1diagram.jpg}
\end{frame}

\begin{frame}{SL-1: Prompt Critical from Manual Control}
  \speak{SL-1 was the first fatal reactor accident in the U.S. One operator withdrew a control rod too far. With no interlocks, the reactor went prompt critical. The resulting steam explosion launched a control rod through the ceiling, killing all three operators \autocite{sl1report}.}
  \begin{itemize}
    \item Root cause: lack of mechanical safeguards
    \item Fallout: Army withdrew from nuclear reactor programs
    \item Image shows post-blast structural damage
  \end{itemize}
\end{frame}

\begin{frame}{Three Mile Island: Interface Confusion}
  \subsubsection*{Misleading Indicators}
  \speak{TMI began with a stuck-open pressure relief valve. A light indicated it had closed. It hadn’t. Operators turned off coolant, thinking the core was overfilled. In truth, the core became exposed and melted \autocite{nrc_tmi_report}.}
  \begin{itemize}
    \item Interface design: misleading feedback lights
    \item Alarm overload: dozens triggered at once
    \item Led to sweeping changes in human-machine interfaces
  \end{itemize}
\end{frame}

\begin{frame}{Three Mile Island: Interface Confusion}
  \centering
  \includegraphics[width=0.75\textwidth]{tmicontrolroom.png}
\end{frame}

\begin{frame}{Chernobyl: Design and Assumption Failures}
  \subsubsection*{When Physics Fights Back}
  \includegraphics[width=0.6\textwidth]{chernobylafter.jpg}
  \speak{This photo shows the wreckage of Chernobyl's Reactor No.~4 just hours after the explosion. The graphite fire continued for days, spewing radiation across Europe \autocite{medvedev1991truth}.}
\end{frame}

\begin{frame}{Chernobyl: Design and Assumption Failures}
  \speak{Chernobyl's RBMK reactor had a positive void coefficient: as coolant boiled, reactivity rose. During a turbine coastdown test, this reactivity led to runaway power. A SCRAM—normally safe—made it worse, as control rods had graphite tips \autocite{iaea1992chernobyl}.}
  \begin{itemize}
    \item Unstable physics design, especially at low power
    \item Operators unaware of full core behavior
    \item Culture of secrecy delayed response and accountability
  \end{itemize}
\end{frame}

\begin{frame}{Fukushima: Nature and Neglect}
  \subsubsection*{Environmental Consequences}
  \includegraphics[width=0.75\textwidth]{fukushimaafter.jpg}
  \speak{Fukushima's tsunami flooded backup generators located in basements. This loss of power—called station blackout—disabled cooling, and all three operating units melted down \autocite{iaea_fukushima_report}.}
\end{frame}

\begin{frame}{Fukushima: Nature and Neglect}
  \speak{Though the reactors shutdown correctly, residual decay heat remained. Without cooling pumps, that heat caused core melt. Desperate engineers tried to wire car batteries together to read instrument panels \autocite{nrc_fukushima}.}
  \begin{itemize}
    \item Failure of backup power design
    \item Hydrogen explosions from vented zirconium reactions
    \item Ocean contamination and public trust fallout
  \end{itemize}
\end{frame}

@@

\begin{frame}{Human Factors Engineering}
  \subsubsection*{Designing for the Operator}
  \speak{Modern control rooms use color-coded touch screens, redundancy, and full-scope training simulators. Human factors seeks to prevent misinterpretation and stress-based mistakes.}
  \begin{itemize}
    \item Cognitive load reduction
    \item Realistic training environments
    \item Alarm prioritization and filtering
  \end{itemize}
\end{frame}

\begin{frame}{Ethics of Automation}
  \subsubsection*{Keeping the Human in the Loop}
  \speak{Machines must serve—not replace—humans in nuclear operations. SCRAMs, alarms, logic interlocks: all must assume a trained operator will intervene. Ethical design means building for fallibility \autocite{mitocw_rbmkreview}.}
  \begin{itemize}
    \item Passive safety aids are essential
    \item But overtrust in automation can fail silently
    \item The human-in-the-loop design is non-negotiable
  \end{itemize}
\end{frame}

@@

\begin{frame}{Aging Infrastructure and Waste}
  \subsubsection*{Legacy Risks}
  \speak{Beyond physics, the nuclear field is strained economically. Most U.S. plants were built pre-1980. Engineering supply chains have shrunk, and no long-term spent fuel site exists \autocite{wna2023chernobyl}.}
  \begin{itemize}
    \item On-site pools reaching capacity
    \item Dry cask storage: temporary fix
    \item Fewer experts entering the field
  \end{itemize}
\end{frame}

\begin{frame}{Modern Developments}
  \subsubsection*{New Hope: SMRs and Vogtle}
  \speak{Still, there is hope. Vogtle Units 3 and 4 in Georgia marked the first new reactors in decades. SMRs and molten salt reactors promise passive safety and easier siting \autocite{worldnuclear_vogtle}.}
  \begin{itemize}
    \item Factory assembly lowers construction risk
    \item Smaller cores, faster SCRAM times
    \item Potential role in green hydrogen and remote grids
  \end{itemize}
\end{frame}

@@

\begin{frame}{Final Thoughts}
  \subsubsection*{Resilience Through Design}
  \speak{From SL-1 to Fukushima, the lesson is clear: safe nuclear power isn't just about neutrons. It's about people, design, and ethics. The atom, when split responsibly, can still light the world.}
  \begin{itemize}
    \item Prioritize human-machine partnerships
    \item Design with humility and feedback
    \item Build public trust through transparency
  \end{itemize}
\end{frame}

\begin{frame}{References}
  \nocite{*}
  \printbibliography
\end{frame}
